
\exercise[(ill-posed boundary value problem)]{2.6}

Consider the following linear boundary value problem
with Dirichlet boundary conditions:
\begin{equation*} 
\begin{split}
&u''(x) +  u(x) = 0 \quad \text{for $a< x< b$}\\
&u(a)=\alpha,\quad u(b)=\beta.
\end{split}
\end{equation*}
Note that this equation arises from a linearized pendulum, for example.

\begin{enumerate} 
\item
Modify the m-file {\tt bvp2.m} to solve this problem.  Test your modified
routine on the problem with
\[
a = 0, \quad b = 1, \quad \alpha = 2, \quad \beta = 3.
\]
Determine the exact solution for comparison.

\item 
Let $a=0$ and $b=\pi$. 
For what values of $\alpha$ and $\beta$  does this boundary value problem
have solutions?
Sketch a family of solutions in a case where there are infinitely many
solutions.

\item Solve the problem with 
\[
a = 0, \quad b = \pi, \quad \alpha = 1, \quad \beta = -1.
\]
using your modified {\tt bvp2.m}.  Which solution to the boundary value
problem does this appear to converge to as $h\goto 0$?  Change the boundary
value at $b=\pi$ to $\beta = 1$.  Now how does the numerical solution behave
as $h\goto 0$?

\item You might expect the linear system in part (c) to be singular since
the boundary value problem is not well posed.  It is not, because of
discretization error.  Compute the eigenvalues of the matrix $A$ for this
problem and show that an eigenvalue approaches 0 as $h\goto 0$.
Also show that $\|A^{-1}\|_2$ blows up as $h\goto 0$ so that the
discretization is unstable.

\end{enumerate} 
